---
title: "A4 Learn Data check"
output: html_notebook
---

This [R Markdown](http://rmarkdown.rstudio.com) Notebook explores the various 
available data for the NOTEPAD project. We want to determine the available
longitudinal data for key biomarkers in cognitively unimpaired data. For this
purpose, we will look at the following biomarkers:
* tau PET
* Amyloid PET
* MRI measurements
* plasma and CSF biomarkers 

The first step is how many unimpaired individuals have followup data in
each of these. We will start with the tables provided in ADNI, and then 
move on to the actual imaging data. 

```{r import}
library(tidyverse)
library(arsenal)
```


Define some helpful globals
```{r globals}
a4_path <- "/Users/davecash/Library/CloudStorage/OneDrive-UniversityCollegeLondon/NOTEPAD/data/A4_LEARN"
```

## Section 1 - TAU PET Data
Load in the Tau PET data and tau QC data
```{r read_dfs}
df_subject <- read_csv(file.path(a4_path,"Derived Data", "SUBJINFO.csv"))
df_amy <- read_csv(file.path(a4_path,"External Data", "imaging_SUVR_amyloid.csv")) %>% 
  filter(brain_region=="Composite_Summary" | brain_region=="lprecuneus_gm") %>% 
  pivot_wider(names_from=brain_region, values_from=suvr_cer) %>% 
  select(-c("suvr_persi","suvr_crus"))
df_tau <- read_csv(file.path(a4_path,"External Data", "imaging_SUVR_tau.csv")) %>% 
  filter(brain_region=="MUBADA Mask" | brain_region=="EarlyTau_VOI") %>% 
  pivot_wider(names_from=brain_region, values_from=c("suvr_persi","suvr_crus")) %>% 
  select(-suvr_cer)

```
``` {r count_tp}
df_tau_group <- group_by(df_tau,BID) %>% 
  mutate(ntp = n()) 
df_tau_bl <- filter(df_tau_group,row_number()==1) %>% ungroup()
count(df_tau_bl,ntp)
tau_substudy<-tableby(SUBSTUDY ~ as.factor(ntp), data=df_tau_bl) 
summary(tau_substudy)

df_amy_group <- group_by(df_amy,BID) %>% 
  mutate(ntp = n()) 
df_amy_bl <- filter(df_amy_group,row_number()==1) %>% ungroup()
count(df_amy_bl,ntp)
amy_substudy<-tableby(SUBSTUDY ~ as.factor(ntp), data=df_amy_bl) 
summary(amy_substudy)

```
